---
title: "R Database Simulations"
format:
  html:
    embed-resources: true
    toc: true
    number-sections: true
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This report demonstrates three R database environments:

1. In-memory database
2. On-disk database
3. Remote database (WRDS)

## Setup
```{r}
library(tidyverse)
library(dbplyr)
library(connections)
library(duckdb)
library(RPostgres)
```

## Create Sample Data

```{r}
stock_prices <- tibble(
  symbol = c(
    rep("AAPL", 6),
    rep("MSFT", 6),
    rep("NVDA", 6)
  ),
  date = rep(seq.Date(as.Date("2024-01-01"), by = "day", length.out = 6), 3),
  adjusted = c(
    180, 181, 179, 183, 185, 186,
    400, 402, 405, 407, 406, 410,
    500, 505, 510, 515, 520, 530
  ),
  volume = c(
    1000000, 1100000, 1200000, 1050000, 980000, 1150000,
    900000, 950000, 1000000, 1020000, 970000, 990000,
    1500000, 1520000, 1490000, 1510000, 1530000, 1550000
  )
)

sp500_index <- tibble(
  symbol = c("AAPL", "MSFT", "NVDA"),
  weight = c(0.07, 0.06, 0.05)
)

stock_prices
sp500_index
```

## 1. In-Memory Database

Open an in-memory DuckDB connection.

```{r}
con <- connection_open(duckdb::duckdb())
con
```

Copy the R data frames into the in-memory database.

```{r}
copy_to(con, stock_prices, name = "stock_price_table", overwrite = TRUE)
copy_to(con, sp500_index, name = "sp500_index_table", overwrite = TRUE)
```

Run a SQL query directly on the database.

```{r}
tbl(
  con,
  sql("
    SELECT symbol, date, adjusted
    FROM stock_price_table
    WHERE symbol = 'AAPL'
    LIMIT 5
  ")
)
```

Assign database tables to R objects for dbplyr.

```{r}
stock_price_db <- tbl(con, "stock_price_table")
sp500_index_db <- tbl(con, "sp500_index_table")

stock_price_db
sp500_index_db
```

Run a dbplyr query.

```{r}
stock_price_db |>
  filter(symbol == "AAPL") |>
  select(symbol, date, adjusted, volume)
```

Run a grouped summary query.

```{r}
stock_price_db |>
  group_by(symbol) |>
  summarize(
    avg_price = mean(adjusted),
    total_volume = sum(volume),
    trading_days = n()
  )
```

Close the in-memory connection.

```{r}
connection_close(con)
```

## 2. On-Disk Database

Create a DuckDB database stored on disk.

```{r}
disk_path <- "disk_db.duckdb"

if (file.exists(disk_path)) {
  unlink(disk_path)
}

con_disk <- connection_open(
  duckdb::duckdb(),
  dbdir = disk_path
)

con_disk
```

Copy the data into permanent tables on disk.

```{r}
copy_to(
  con_disk,
  stock_prices,
  name = "stock_price_table_disk",
  temporary = FALSE,
  overwrite = TRUE
)

copy_to(
  con_disk,
  sp500_index,
  name = "sp500_index_table_disk",
  temporary = FALSE,
  overwrite = TRUE
)
```

Query the on-disk tables.

```{r}
stock_price_db_disk <- tbl(con_disk, "stock_price_table_disk")
sp500_index_db_disk <- tbl(con_disk, "sp500_index_table_disk")

stock_price_db_disk |>
  filter(adjusted > 200) |>
  select(symbol, date, adjusted, volume)
```

Show that the database file exists on disk.

```{r}
file.info(disk_path)[, c("size", "mtime")]
```

Close and reopen the connection to prove persistence.

```{r}
connection_close(con_disk)

con_disk2 <- connection_open(
  duckdb::duckdb(),
  dbdir = disk_path
)

con_disk2
```

Run another query after reconnecting.

```{r}
tbl(con_disk2, "stock_price_table_disk") |>
  group_by(symbol) |>
  summarize(
    avg_price = mean(adjusted),
    total_volume = sum(volume),
    trading_days = n()
  )
```

Close the on-disk connection.

```{r}
connection_close(con_disk2)
```

## 3. Remote Database (WRDS)

This section connects to WRDS through PostgreSQL.

Set your WRDS username and password in the R console before rendering.

```{r}
# Example only:
# Sys.setenv(WRDS_USER = "your_username")
# Sys.setenv(WRDS_PASSWORD = "your_password")
```

Check whether credentials are available.

```{r}
#| include: false

Sys.setenv(
  WRDS_USER = "jean124",
  WRDS_PASSWORD = "!bF#3jf&ZX2ftW?"
)

wrds_user <- Sys.getenv("WRDS_USER")
wrds_password <- Sys.getenv("WRDS_PASSWORD")
wrds_ready <- wrds_user != "" && wrds_password != ""
```

Open the WRDS connection only if credentials exist.

```{r}
if (wrds_ready) {
  wrds_con <- connection_open(
    RPostgres::Postgres(),
    host = "wrds-pgdata.wharton.upenn.edu",
    port = 9737,
    dbname = "wrds",
    sslmode = "require",
    user = wrds_user,
    password = wrds_password
  )
  
  cat("Connected to WRDS successfully.\n")
} else {
  cat("WRDS credentials not found. Set WRDS_USER and WRDS_PASSWORD, then render again.\n")
}
```

Browse schemas/tables on WRDS.

```{r}
if (wrds_ready) {
  all_schema <- tbl(wrds_con, I("information_schema.tables")) |>
    filter(table_type %in% c("VIEW", "FOREIGN TABLE")) |>
    distinct(table_schema, table_name) |>
    collect()

  head(all_schema, 10)
}
```

Optionally inspect one schema shown in your slides.

```{r}
if (wrds_ready) {
  all_schema |>
    filter(table_schema == "fisd") |>
    head()
}
```

Access Compustat fundamentals remotely.

```{r}
if (wrds_ready) {
  funda_db <- tbl(wrds_con, I("comp.funda"))
  funda_db
}
```

Run a remote query and collect results into R.

```{r}
if (wrds_ready) {
  funda_sample <- funda_db |>
    filter(
      indfmt == "INDL",
      datafmt == "STD",
      consol == "C",
      datadate >= as.Date("1990-01-01")
    ) |>
    select(gvkey, datadate, seq, ceq, at, lt, capx, oancf) |>
    head(10) |>
    collect()

  funda_sample
}
```

Save the result to CSV and show the file size.

```{r}
if (wrds_ready) {
  write_csv(funda_sample, "wrds_funda_sample.csv")
  file.info("wrds_funda_sample.csv")[, "size", drop = FALSE]
}
```

Close the WRDS connection.

```{r}
if (wrds_ready) {
  connection_close(wrds_con)
}
```

## Conclusion

This report demonstrated:

- an in-memory DuckDB database
- an on-disk DuckDB database
- a remote WRDS PostgreSQL database

The HTML output includes both code and visible results, and `embed-resources: true` makes the rendered file self-contained.